---
title: "The {healthyverse}"
subtitle: "A Presentation to the SoCal RUG"
author: "Steven P Sanderson II, MPH"
date: "`r Sys.Date()`"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## About Me

Name - Steven P Sanderson II, MPH

Education:

- BA Economics: SUNY Stony Brook University
- MPH: SUNY Stony Brook University

Employment:

- Data Scientist/IT Manager: NYU Langone Healthy at Long Island Community Hospital
- Head Data Scientist: Manchu Technologies Corp. (Consulting in SQL, R)

## Sites

GitHub: https://github.com/spsanderson

LinkedIn: https://www.linkedin.com/in/spsanderson/

Personal: https://www.spsanderson.com

## Interests

- Data Analysis
- R (probably obvious)
- Plex Server (Music is life) - Is 88K mp3's to much? NO!

## How I got Started

I got started with R in my first job out of undergrad.

Employer: "Do you know SQL and any other analysis programming?"
Me: Yes, not a problem at all.

Reality...a little different, maybe an alternative fact.

SO...Google Searching...saved the day? YES!

## Slide with Bullets

- Bullet 1
- Bullet 2
- Bullet 3

## Slide with R Output

```{r cars, echo = TRUE}
summary(cars)
```

## Slide with Plot

```{r pressure}
plot(pressure)
```

