---
title: "The {healthyverse}"
subtitle: "A Presentation to the SoCal RUG"
author: "Steven P Sanderson II, MPH"
date: "`r Sys.Date()`"
output: 
    ioslides_presentation:
        logo: healthyverse.png
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(tidyquant)
library(timetk)
library(healthyverse)
library(tidymodels)
library(modeltime)
```

## About Me

Name - Steven P Sanderson II, MPH

Education:

- BA Economics: SUNY Stony Brook University
- MPH: SUNY Stony Brook University

Employment:

- Data Scientist/IT Manager: NYU Langone Health at Long Island Community Hospital
- Head Data Scientist: Manchu Technologies Corp. (Consulting in `SQL`, `R`)

Developer:

- `healthyverse` packages.

## Sites

GitHub: https://github.com/spsanderson

LinkedIn: https://www.linkedin.com/in/spsanderson/

Personal: https://www.spsanderson.com

## Interests

- Data Analysis
- R (probably obvious)
- Plex Server (Music is life) - Is 88K mp3's to much? NO!

## How I got Started

I got started with R in my first job out of undergrad.

Employer: "Do you know SQL and any other analysis programming?"
Me: Yes, not a problem at all.

Reality...a little different, maybe an alternative fact.

SO...Google Searching...saved the day? YES!

## What are we going to talk about

So what are we going to discuss today?

The `healthyverse` suite of packages.

## Healthyverse Packages

There are currently five packages in the `healthyverse`


| Package | r-universe badge |
|:-------:|:----------------:|
| `healthyR` |  [![healthyR status badge](https://spsanderson.r-universe.dev/badges/healthyR)](https://spsanderson.r-universe.dev) |
| `healthyR.data` |  [![healthyR.data status badge](https://spsanderson.r-universe.dev/badges/healthyR.data)](https://spsanderson.r-universe.dev) |
| `healthyR.ai` | [![healthyR.ai status badge](https://spsanderson.r-universe.dev/badges/healthyR.ai)](https://spsanderson.r-universe.dev) |
| `healthyR.ts` | [![healthyR.ts status badge](https://spsanderson.r-universe.dev/badges/healthyR.ts)](https://spsanderson.r-universe.dev) |
| `TidyDensity` | [![TidyDensity status badge](https://spsanderson.r-universe.dev/badges/TidyDensity)](https://spsanderson.r-universe.dev) |

The `healthyverse` package is really just a meta package like tidyverse.

## The Begining

First came `healthyR`

The desire of this packge is to be...

It is experimental

[Site: www.spsanderson.com/healthyR](www.spsanderson.com/healthyR)

## `healthyR`

Current Stats

- Has some functionality that can be useful right out of the gate
- Plotting functions like `los_ra_index_plt()`, `kmeans_scree_plt()`, and `gartner_magic_chart_plt()`

## LOS RA Index Plot
This plot takes place at the end of a work flow. Let's take a look at an example. First we need some data.

```{r warning=FALSE, message=FALSE}
set.seed(123)
data_tbl <- tibble(
  "alos"                 = rnorm(186, 6.5, 3)
  , "elos"               = rnorm(186, 6, 3)
  , "readmit_rate"       = rnorm(186, .1, .05)
  , "readmit_rate_bench" = rnorm(186, .075, .05)
)

data_tbl
```

## LOS RA Index Plot (summary)
Now that we have our data, let's take a look at the summary table.

```{r warning=FALSE, message=FALSE}
summary_tbl <- los_ra_index_summary_tbl(
  .data = data_tbl
  , .max_los       = 12
  , .alos_col      = alos
  , .elos_col      = elos
  , .readmit_rate  = readmit_rate
  , .readmit_bench = readmit_rate_bench
)

head(summary_tbl, 5)
```

## LOS RA Index Plot (plot)

Now lets take a look at the resulting plot and discuss what it tells us.

```{r warning=FALSE, message=FALSE}
los_ra_index_plt(summary_tbl)
```

## K-Means

What is it? In it's simplest form, K-Means will partition data into K non-overlapping clusters.

Example? Think a bank wanting to give people credit cards, this is a clustering
problem, high income, average income, low income, etc.

It's easy to understand how many applications this has.

Healthcare is no different.

## Scree Plot

We are only going to focus on the scree plot. This helps us to decide how many
clusters are in our data.

This plot is also many times referred to as the elbow plot.

## Example

```{r warning=FALSE, message=FALSE}
data_tbl <- healthyR_data%>%
   filter(ip_op_flag == "I") %>%
   filter(payer_grouping != "Medicare B") %>%
   filter(payer_grouping != "?") %>%
   select(service_line, payer_grouping) %>%
   mutate(record = 1) %>%
   as_tibble()

ui_tbl <-  kmeans_user_item_tbl(
   .data           = data_tbl
   , .row_input    = service_line
   , .col_input    =  payer_grouping
   , .record_input = record
 )

kmm_tbl <- kmeans_mapped_tbl(ui_tbl)

kmeans_scree_plt(.data = kmm_tbl)
```

## What Just Happened?

What just happened? Magic? Maybe?

First the data. Using the `healthyR_data` from my package `healthyR.data` we get a tibble of records with a service line and a payer grouping.

A patients service line is the row record and the payer grouping is the item, this gives us the user item matrix we need.

Next we get a tibble with multiple centers via the function `kmeans_mapped_tbl()`

Lastly, we plot that mapped tibble.

## Magic Chart

Have you seen the Gartner Magic Quadrant Chart?

If so, then you know, if not, then I'll show.

First let's take a look at the function call.

## Function Call

```{r, eval=FALSE, echo=TRUE}
gartner_magic_chart_plt(
  .data = tibble::tibble(
      x = rnorm(100, 0, 1)
      , y = rnorm(100,0,1)
  )
  , .x_col = x
  , .y_col = y
  , .x_lab = "LOS"
  , .y_lab = "Readmit"
  , .plt_title = "Test Plot"
  , .tr_lbl = "High RA-LOS"
  , .tl_lbl = "High RA"
  , .bl_lbl = "Leader"
  , .br_lbl = "High LOS"
)
```

## Plot Output

```{r}
gartner_magic_chart_plt(
  .data = tibble::tibble(
      x = rnorm(100, 0, 1)
      , y = rnorm(100,0,1)
  )
  , .x_col = x
  , .y_col = y
  , .x_lab = "LOS"
  , .y_lab = "Readmit"
  , .plt_title = "Test Plot"
  , .tr_lbl = "High RA-LOS"
  , .tl_lbl = "High RA"
  , .bl_lbl = "Leader"
  , .br_lbl = "High LOS"
)
```

## `healthyR.ts`

What is `healthyR.ts`

[Site: www.spsanderson.com/healthyR.ts](www.spsanderson.com/healthyR.ts)

The intent is that it be a easy to use verb framework for quick modeling of
time-series data.

It is experimental.

In this section we are going to go over the processes of model tuning.

## Time Series Data

For this we use the internal `AirPassengers` data set.

```{r, echo=TRUE}
data_ts <- AirPassengers

head(data_ts)
```

## Transform to `tibble` 

Most of the work done inside of `healthyR.ts` requires the use of a `tibble`, and where a `ts` object can be used, it is typically converted to a `tibble` internally to the function. This is done on purpose so that it works well with the rest of the `tidyverse`

```{r echo=TRUE}
data_tbl <- ts_to_tbl(data_ts) %>%
  select(-index)

head(data_tbl, 5)
```

## `ts` to `tibble`

With the function `ts_to_tbl()` we can take a time-series object and make it into a `tibble` object.

## Splits Oject

Now that we have our data, we need to make our time series split object. We use `timetk::time_series_split()`

```{r, echo=TRUE}
splits <- time_series_split(
    data_tbl
    , date_col
    , assess = 12
    , skip = 3
    , cumulative = TRUE
)

head(training(splits), 1)
```

## Recipe

Now that we have our splits object we can go ahead and make the recipe. For this
we will use `ts_auto_recipe()` which creates four different recipes and outputs them into a list object.

```{r, echo=TRUE}
rec_objs <- ts_auto_recipe(
  .data       = data_tbl
  , .date_col = date_col
  , .pred_col = value
)

```

## Recipe (con't)

Lets take a look at it

```{r, echo=TRUE}
rec_objs[[1]]
```

We are not going to look at all of them.

## Workflow Sets and `healthyR.ts`
Now that we have our data and our recipe, we can make our automatic `workflowsets` object. As mentioned at the top we are going to use the MARS algorithm with the earth engine. This function from `healthyR.ts` will take in a list of recipes and a `.model_type parameter`, and create a `workflowsets` `tibble`, where `cross` is set to `TRUE`

We use `ts_wfs_mars()` in this example.

```{r eval=FALSE, echo=TRUE}
ts_wfs_mars(
  .model_type = "earth",
  .recipe_list,
  .num_terms = 200,
  .prod_degree = 1,
  .prune_method = "backward"
)
```

## (con't)

Lets make the workflowset objec.

```{r echo=TRUE}
wfsets <- healthyR.ts::ts_wfs_mars(
  .model_type = "earth"
  , .recipe_list = rec_objs
)

wfsets
```

## (con't)

I tried to give reasonable defaults for all of the functions used in the `healthyverse` set of packages.

Many times hospitals, especially smaller non-academic or even small academic ones do not have a department of analysts. It could be that there are none or just one.

## Modeltime workflow

This workflow is outside of the scope of the `healthyverse` but necessary as it is an intermediate step before we get to the tuning process.

```{r echo=TRUE}
wf_fits <- wfsets %>%
  modeltime_fit_workflowset(
    data = training(splits)
    , control = control_fit_workflowset(
     allow_par = FALSE
     , verbose = FALSE
    )
  )

models_tbl <- wf_fits %>%
  filter(.model != "NULL")
```

## (con't)

You will see that we did `.model != "NULL"` this is because if a certain model tries to run, lets say an `ets` model where alpha < beta then it will fail and the `.model` output will be equal to `"NULL"`

Lets see the models_tbl

```{r echo=TRUE}
models_tbl
```

## Calibration 

This is another part of the `modeltime` workflow that is used. We _calibrate_ the data with the testing split using `testing(splits)`

```{r echo=TRUE}
calibration_tbl <- models_tbl %>%
  modeltime_calibrate(new_data = testing(splits))

calibration_tbl
```

## Automatic Tuning

```{r eval=FALSE, echo=TRUE}
output <- healthyR.ts::ts_model_auto_tune(
  .modeltime_model_id = 1,
  .calibration_tbl    = calibration_tbl,
  .splits_obj         = splits,
  .drop_training_na   = TRUE,
  .date_col           = date_col,
  .value_col          = value,
  .tscv_assess        = "12 months",
  .tscv_skip          = "3 months",
  .num_cores          = 1
)
```

## What happened?

We just automatically tuned the parameters of a model.

## Outputs

```{r echo=FALSE, message=FALSE, warning=FALSE}
output <- healthyR.ts::ts_model_auto_tune(
  .modeltime_model_id = 1,
  .calibration_tbl    = calibration_tbl,
  .splits_obj         = splits,
  .drop_training_na   = TRUE,
  .date_col           = date_col,
  .value_col          = value,
  .tscv_assess        = "12 months",
  .tscv_skip          = "3 months",
  .num_cores          = 5
)
```

There are three sections of output from the function.

1. Data
2. Model Info
3. Plots

## Data

The function returns the following:

1. Calibration tibble from modeltime workflow
2. Calibration tuned tibble from tuned model
3. Time Series Cross Validation tibble
4. Tuned Model Results
5. Best tuned results
6. Time Series Cross Validation object (rsample object)

## Model Info

The following model information is available:

1. Model Specification
2. Parsnip Engine used
3. Plucked Model (The original model before tuning)
4. Tuned Workflow specification
5. Tuning Grid used
6. Tuned Time Series Cross Validation Workflow 

## Plots

The following plots are returned;

1. Tuning Grid Plot
2. Cross Validation Plan

https://www.spsanderson.com/healthyR.ts/articles/auto-model-tune.html#data


## `healthyR.ai`

What is `healthyR.ai`?

The desire of this packge is to be...

It is experimental

[Site: www.spsanderson.com/healthyR.ai](www.spsanderson.com/healthyR.ai)

## `healthyR.data`

## `TidyDensity`